Simulating High-Dimensional Multivariate Data

### 2020-11-06 050-examples.Rmd
2nd draft. scale up d in quad loss, drop m to 10

## Simulation-based UHD joint probability calculations

@@

```{r ch010-myM, include=FALSE, cache = TRUE}
m <- 100
```

MC methods are routinely used in many statistical inferential tasks including estimation, hypothesis testing, error rates, and empirical interval coverage rates.
For an concise introduction to these methods, see, for example @Rizzo2007, Ch. 6.
Now, we demonstrate how `bigsimr` can be used evaluate estimation efficiency.
In particular, we'd like to assess the error our correlation estimation above.
We used a conventional method, based on classical statistical theory. 
Yet this method was not designed for high-dimensional data.
Indeed, high-dimensional covariance estimation (and precision matrices) is an active area of statistical science (see, for example, @Won2013g and @VanWieringen2016).
In this example, we simulate $m=`r m`$ data sets with the number of simulated vectors matching the number of patients in the BRCA data set, $N=`r nrow(brca)`$.
At each iteration we estimate the quadratic loss from the specified $\bf{R^{Spearman}_Y}$ and average over the $m$ iteration.

```{r ch050-quadloss, echo=TRUE, eval=FALSE, cache=FALSE}
quadLossRDS <- paste0( "results/brca", round(myProb*100), "quadLoss.rds" )
if ( !file.exists( quadLossRDS) )  {
    n <- nrow(brca)
    simRho <- replicate(n = m, expr = { tmpSim <- rvec(n = nrow(brca) , nb_Rho, nb_margins,
                                                       type = corType, ensure_PSD = TRUE, cores = cores)
                                                       tmpRho <- bigsimr::cor_fast( x = tmpSim )} ,
                        simplify = FALSE)

    quadLoss <- unlist( lapply( simRho, rags2ridges::loss, T = nb_Rho, type = "quadratic"  ) )
    saveRDS(quadLoss, quadLossRDS)
}
quadLoss <- readRDS( quadLossRDS )
```

@@

```{r ch050-quadloss-summary, echo=TRUE, eval=FALSE, cache=FALSE}
summary(quadLoss)
```

>>>>

## Simulation-based joint probability calculations

@@

```{r ch010-myM, include=FALSE, cache = TRUE}
m <- 10
```

MC methods are routinely used in many statistical inferential tasks including estimation, hypothesis testing, error rates, and empirical interval coverage rates.
For an concise introduction to these methods, see, for example @Rizzo2007, Ch. 6.
Now, we demonstrate how `bigsimr` can be used evaluate estimation efficiency.
In particular, we'd like to assess the error our correlation estimation above.
We used a conventional method, based on classical statistical theory. 
Yet this method was not designed for high-dimensional data.
Indeed, high-dimensional covariance estimation (and precision matrices) is an active area of statistical science (see, for example, @Won2013g and @VanWieringen2016).
In this example, we simulate $m=`r m`$ data sets with the number of simulated vectors matching the number of patients in the BRCA data set, $N=`r nrow(brca)`$.
Using the fact that our simulation is much faster for the Pearson correlation type (see Figure \@ref(fig:ch040-gpuVScpuFig), we convert the Spearman correlation matrix and ensure PSD outside the `rvec` call.
At each iteration, we estimate the quadratic loss from the specified ${\bf R}^{Spearman}_{ \bf Y}$, producing a distribution of loss values.

```{r ch050-quadloss, echo=TRUE, eval=T, cache=FALSE}
quadLossRDS <- paste0( "results/brca", round(myProb*100), "quadLoss.rds" )
if ( !file.exists( quadLossRDS) )  {
    n <- nrow(brca)
    ## convert outside for faster simulation
    nb_Rho_p <- bigsimr::cor_convert( rho = nb_Rho, from = corType, to = "pearson" )
    ## ensure PSD
    nb_Rho_p <- bigsimr::cor_nearPSD( G = nb_Rho_p )
    ## create m random vectors and estimate correlation
    simRho <- replicate(n = m,
                        expr = { tmpSim <- rvec(n = n , nb_Rho_p, nb_margins,
                                                type = 'pearson', ensure_PSD = FALSE, cores = cores);
                                                bigsimr::cor_fast( x = tmpSim, method = corType )} ,
                        simplify = FALSE)
    ## find quadratic loss at each rep
    quadLoss <- unlist( lapply( simRho, rags2ridges::loss, T = nb_Rho, type = "quadratic"  ) )
    saveRDS(quadLoss, quadLossRDS)
}
quadLoss <- readRDS( quadLossRDS )
```

@@

```{r ch050-quadloss-summary, echo=TRUE, eval=T, cache=FALSE}
summary(quadLoss)
```
